Computing Entropies With Nested Sampling

### 2017-08-14 paper/response.tex

That's why I started thinking about this problem...

{\quote
1. The author presents a methodology for computing entropies, and other information-theoretic quantities, using Skilling's nested sampling algorithm.  It should be noted that Skilling himself advocates the estimation of the "entropy", I believe that it is the KL-divergence, as the estimate of the uncertainty in the estimated evidence.  This should be mentioned, and possibly discussed within the context of this present paper.}

{\quote
2. With regard to the algorithm to compute the depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.  Also at that step, do the square brackets refer to an array (of one element)?  Or do they refer to some kind of rounding, which I doubt.}

Sorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
\end{align}
which is hopefully more understandable.

@@

{\quote
7. After (26) what is meant by "NS repetitions"?}

It's the number of reference particles
whose depth was estimated using an NS run. I've added a sentence to this effect.

{\quote
8. After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"}

{\quote
9. It wasn't clear to me how the distance function (25) was chosen and why.}

It is just the Euclidean distance between two data vectors, each of which
contains 100 co-ordinates. I added a sentence to the paper to clarify this.

{\quote
10. I had some questions about how the algorithm was conceived.  I had a difficult time envisioning how the blue distribution played the role of the prior, and why.  Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.}

I know a few groups use that convention (transform the parameter space so
the prior becomes uniform), but I never have, and there's no reason to do
that in general, as long as it's possible to explore $\pi$ then everything
should be fine.

{\quote
11.  It was not entirely clear to me how the algorithm was used to estimate the entropy of a Gaussian.  How was the posterior (24) arrived at.  The prior is normal.  But what about the likelihood?  And then how was the distance measure (25) computed?}

I have added an extra paragraph with the intent of explaining this in more
detail.

### 2017-08-14 paper/ms.tex
More work on Kevin's comments

Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\boldsymbol{x} = \{x_1, ..., x_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
\begin{align}
\mu | \boldsymbol{x} &\sim \textnormal{Normal}\left(
                                       \frac{1}{100.01}\sum_{i=1}^{100} x_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
The true value of the differential entropy $H(\boldsymbol{x})$ is
available analytically
in this case, and its value is $146.499$ nats. However, in most situations, the
entropy of a marginal distribution is not available in closed form.

To use the algorithm, a distance function $d(;)$ must be defined, to quantify
how far the NS particles are from the reference particle. To calculate
the entropy of the data, I used the following distance function,
{\revision based on
a simple Euclidean metric in the space of datasets:
\begin{align}
d\left(\x; \xref\right) &= \sqrt{
        \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2}
\end{align}
}

>>>>

Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\boldsymbol{x} = \{x_1, ..., x_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(x_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
{\revision
I assumed a Normal$(0, 10^2)$ prior for $\mu$.
Before getting the data, the prior information is described by the joint
distribution
\begin{align}
p(\mu, \boldsymbol{x}) &= p(\mu)p(\boldsymbol{x} | \mu).\label{eqn:joint_prior}
\end{align}
The prior over datasets is given by the marginal distribution
\begin{align}
p(\boldsymbol{x}) &= \int p(\mu, \boldsymbol{x}) \, d\mu.
\end{align}
This is the distribution whose differential entropy I seek in this section.
Its true value is
available analytically
in this case, by noting that the posterior is available in closed form
for any dataset:
\begin{align}
\mu | \boldsymbol{x} &\sim \textnormal{Normal}\left(
                                       \frac{1}{100.01}\sum_{i=1}^{100} x_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
which enables the calculation of the mutual information and hence
$H(\boldsymbol{x})$.
The true value of $H(\boldsymbol{x})$ is $146.499$ nats.
However, in most situations, the
entropy of a marginal distribution such as this is not available in closed form.

To use the algorithm, a distance function $d(;)$ must be defined, to quantify
how far the NS particles are from the reference particle. To calculate
the entropy of the data, {\revision I applied the algorithm in the joint
space of possible parameters and datasets}, i.e., distribution of reference
particles followed Equation~\ref{eqn:joint_prior}.
For the distance function, which helps the NS particles approach the
reference particle, I used a
simple Euclidean metric in the space of datasets:
\begin{align}
d\left(\x; \xref\right) &= \sqrt{
        \sum_{i=1}^{100}\left(x_i - x_{{\rm ref}, i}\right)^2}.
\end{align}
Since the distance function refers only to the data $\boldsymbol{x}$
and not the parameter $\mu$, the sampling is effectively done in the
space of possible datasets only --- the parameter $\mu$ can be thought of
as merely a
latent variable that allows $p(\boldsymbol{x})$ to be explored.
}
